### 2020-04-15 experiments.tex
ok

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We also use these results to answer a set of research questions that consider the utility of our method:

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
\item {\bf RQ5:}  Are killed mutants reasonablely similar to actual faults?  That is, is the data above based on meaningful changes to code?
\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?
\item {\bf RQ7:}  How easy is it to identify low-hanging fruit for improving tools \emph{without} prioritization to remove conceptually redundant mutants?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a more precise and accurate tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.  The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.  Chekam et al. showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.  {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.  However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.  This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
{\bf RQ6} and {\bf RQ7} receive only preliminary answers, for the smart contract tools.  We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible.

@@

\subsubsection{Improving Slither {\bf (RQ6 and RQ7)}}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that for {\bf RQ7} it is indeed hard to use mutation analysis results without prioritization. A large majority of the mutants we inspected involved either the missing
{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior. SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6} and {\bf RQ7}.

@@

We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.  This tool has been heavily tuned and used for for Solidity and Python, and moderately tuned for Java.  The Java mutation also benefits from extensive tuning on C programs, which share syntactic mutation patterns with both Java and Solidity.
There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers.  We will de-anonymize this if the paper is accepted.

>>>>

Our primary experimental results are a set of comparisons of tools using our method, for three languages: Solidity (the most popular language for smart contracts), Java, and Python.  We used the Universal Mutator tool for all experiments, since e.g., PIT \cite{pitmut} does not produce source-level mutants and the PMD tool requires source code; we also wanted to use a consistent mutation generation method for all languages, to the extent possible; for Solidity and Python, we believe the Universal Mutator is simply the best available tool.   We also our results to answer a set of research questions that consider the utility of our method:

\begin{itemize}
\item {\bf RQ1:}  Does mutation analysis of static analysis tools produce actionable results?  That is, do raw mutation kills serve to distinguish tools from each other, or are all tools similar in terms of mutation detection capability?
\item {\bf RQ2:}  Does our approach provide additional information beyond simply counting findings for the original, un-mutated analyzed code?  Do \emph{ratios} differ between programs, or does the number of mutants killed simply reflect the ``verbosity'' of each tool?
\item {\bf RQ3:}  Do the rankings that raw kills and ratios establish agree with other sources of information about the effectiveness of the evaluated tools?  Can we confirm results from other evaluation methods, or informal opinion?
\item {\bf RQ4:}  Are mutation scores for clean programs higher (more mutants killed) than those for programs for which tools report findings before mutation?
\item {\bf RQ5:}  Are killed mutants reasonablely similar to actual faults?  That is, is the data above based on meaningful changes to code?
\item {\bf RQ6:}  Do individual mutants, prioritized for ease of examination, allow us to identify classes of faults that
 different tools are good at/bad at, and extend this information to improve tools by addressing identified weaknesses?  How does this compare to using mutants that have not been prioritized?
  \end{itemize}

In particular, we consider {\bf RQ2} to be of critical importance; if the mutant ratios for tools differ, then this is clear evidence that our hypothesis that the tendency of mutants to be faults, and to expect that mutated code will, by a more precise and accurate tool, be flagged as problematic more often than non-mutated code, holds.  This expectation that (some subset of the) mutants can serve as proxies for real, detectable faults is the core concept of our approach.
{\bf RQ4} addresses a concern briefly mentioned in the introduction:  it is possible that warnings for the original code interfere with our definition of detection.  The ideal case for our approach is when a tool report no findings for un-mutated code, and either reports a finding (kills the mutant) or not (does not kill it) when the mutant is introduced.  Chekam et al. showed that the ``clean program assumption'' for testing, the idea that coverage will be similar for faulty and fixed versions of a program, is a threat to the validity of investigations of the relationship between coverage and fault detection \cite{CleanProgram}, and we want to establish that this is unlikely to be the case for our approach.  {\bf RQ5} is somewhat inherently qualitative and incomplete; we cannot analyze all mutants on which results are based manually, and understanding the mutants and tool warnings completely would require deep understanding of all the software systems investigated.  However, in many cases, the impact of a mutant is clear, and the reason for warnings is obvious.  This was often enough the case that, as we discuss below, we are confident mutants that distinguish tools are meaningful (missed) opportunities for static analysis tools.
For {\bf RQ6}  we have only preliminary answers, for the smart contract tools.

@@

\subsubsection{Improving Slither {\bf (RQ6)}}

@@

Examining the first 100 mutants in the unprioritized lists for SmartCheck and Securify, ordered by contract ID and mutant number (roughly source line mutated) we were unable to identify \emph{any} obviously interesting mutants, suggesting that it is indeed hard to use mutation analysis results without prioritization. A large majority of the mutants we inspected involved either the missing
{\tt return} problem noted in the introduction, or replacing {\tt msg.sender} with {\tt tx.origin}; Slither, of course, has a detector for misuses of {\tt tx.origin}, and we believe (but are not sure) that almost all of the differences with respect to such changes are due to intentional behavior. SmartCheck and Securify tend to identify most (though not all) uses of {\tt tx.origin} as incorrect, while Slither has a more selective rule, intended to reduce false positives.
It is hard to scale our efforts here to a larger experiment, since writing and submitting changes to static analysis tools is always going to be a fairly onerous task, but we believe that our successful addition of new detectors, and the ease of identifying good candidate detectors using mutant prioritization supports a limited affirmative answer to {\bf RQ6}.

@@

We used the Universal Mutator \cite{universalmutator,regexpMut}, which aggressively produces large numbers of mutants, but does not target any particular software defect patterns, to generate all mutants.
There is no room in the paper to present the exact set of projects analyzed, but we have provided an (anonymized) github repository containing raw results for inspection by reviewers, or further analysis by other researchers (\url{the repo}).

### 2020-04-15 introduction.tex
ok

Static analysis of code is one of the most effective ways to avoid defects in software, and when security is a concern, applying effective static analysis tools is essential.  Static analysis can find problems that are extremely hard to detect by testing, e.g. when the inputs triggering a bug are hard to find.  Static analysis is also often more efficient than testing; a bug that takes a fuzzer like AFL days to find may be immediately identified by a good static analysis tool.

@@

Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, most small changes will therefore introduce a fault.  For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.  Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
In an ASE 2015 \cite{groce2015verified} paper and a 2018 journal extension \cite{groce2018verified} of that paper, Groce et al. proposed examining individual mutants that survive a formal verification or automated test generation process to detect and correct weaknesses in a specification or test generator.  The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and improve a heavily used test generator for the {\tt pyfakefs} file system, allowing it to detect multiple new bugs.  Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale to large code bodies.

>>>>

Static analysis of code is one of the most effective ways to avoid defects in software, and, when security is a concern, is simply essential.  Static analysis can find problems that are extremely hard to detect by testing, when the inputs triggering a bug are hard to find.  Static analysis is also often more efficient than testing; a bug that takes a fuzzer like AFL days to find may be immediately identified by a good static analysis tool.

@@

Mutation testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical} (or mutation analysis, the term we will use in this work, for reasons that will become clear) uses small syntactic changes to a program to introduce synthetic ``faults,'' under the assumption that if the original version of a program is mostly correct, most small changes will therefore introduce a fault.  For the most part, mutation analysis has been used to evaluate test suites by computing a mutation score, the fraction of mutants the suite detects, or ``kills''.  Most such use has been in research efforts, rather than practical testing efforts, though there has been sporadic use by interested developers.
Groce et al. \cite{groce2015verified,groce2018verified} proposed examining individual mutants that survive a rigorous testing and verification effort to detect and correct weaknesses in such efforts.  The approach was able to expose bugs in a heavily-tested module of the Linux kernel \cite{mutKernel} and a widely used Python file system.  Recently, mutation analysis has been adopted in industrial settings, though not for actual examination of all surviving mutants \cite{MutGoogle,ivankovic2018industrial}, a practice that is hard to scale.
